Lapse & Vision Loss PS Analysis

### 2025-03-26 analysis.Rmd
save diagnostics

      love_plot <- if (params$method != "twang") { 
        tryCatch({
          love.plot(result, stat = "mean.diffs", var.order = "adjusted", stars = "std")
        }, error = function(e) {
          message("Error generating balance plot for ", params$name, ": ", e)
          return(NULL)
        })
      } else {
        NULL
      }

@@

@@

      saveRDS(list(love_plot = love_plot, balance_plot = balance_plot, summary = balance_summary), file = diagnostics_filename)
    } else {
      print("Loading cached diagnostics...")
      diagnostics <- readRDS(diagnostics_filename)
      love_plot <- diagnostics$love_plot
      balance_plot <- diagnostics$balance_plot
      balance_summary <- diagnostics$summary
    }

    all_results[[params$name]] <- list(
      result = result,
      love_plot = love_plot,
      balance_plot = balance_plot,
      summary = balance_summary
    )
  }

@@

# Define a list of matching methods and parameters, each with a unique name
methods <- list(
  list(name = "nearest_glm", method = "nearest", distance = "glm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_gam", method = "nearest", distance = "gam", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_gbm", method = "nearest", distance = "gbm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_lasso", method = "nearest", distance = "lasso", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_rpart", method = "nearest", distance = "rpart", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "subclass_glm", method = "subclass", distance = "glm", subclasses = 5, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "nearest_mahalanobis", method = "nearest", distance = "mahalanobis", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE),
  list(name = "cbps", use="weightit", method="cbps", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "entropy", use="weightit", method="ebal", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "bart", use="weightit", method="bart", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = FALSE),
  list(name = "twang_gbm", use="twanng", method = "twang", estimand="ATT", rerun_model = FALSE, rerun_diagnostics = FALSE)
)

@@

```{r love_plots}
for (method_name in method_names) {
  # Get balance and common support plots for the method
  method_results <- results[[method_name]]
  love_plot <- method_results$love_plot
  balance_plot <- method_results$balance_plot

@@

  # Output section header for balanace plot of pscore (if exists)
  if (!is.null(balance_plot)) {
    cat("### Balance Plot of distance measure for", method_name, "\n\n")
    print(balance_plot)
    cat("\n\n")
  }
}
```

>>>>

      love_plot <- if (params$method != "twang") { 
        tryCatch({
          love.plot(result, stat = "mean.diffs", var.order = "adjusted", stars = "std")
        }, error = function(e) {
          message("Error generating balance plot for ", params$name, ": ", e)
          return(NULL)
        })
      } else if(params$method == "twang") {
       tryCatch({
          plot(result, plots=3)
        }, error = function(e) {
          message("Error generating balance plot for ", params$name, ": ", e)
          return(NULL)
        })
      } else {
        NULL
      }

@@

      p_dist_plot <- tryCatch({
        if (params$method == "twang") {
          plot(result, plots=2)
        } else if("distance" %in% names(result)) {
          plot(result, type = "jitter", interactive=FALSE)
        } else {
          NULL
        }
      }, error = function(e) {
        message("Error generating p-score distribution support plot for ", params$name, ": ", e)
        return(NULL)
      })

@@

      saveRDS(list(love_plot = love_plot, balance_plot = balance_plot, p_dist_plot = p_dist_plot, summary = balance_summary), file = diagnostics_filename)
    } else {
      print("Loading cached diagnostics...")
      diagnostics <- readRDS(diagnostics_filename)
      love_plot <- diagnostics$love_plot
      balance_plot <- diagnostics$balance_plot
      balance_summary <- diagnostics$summary
    }

    all_results[[params$name]] <- list(
      result = result,
      love_plot = love_plot,
      p_dist_plot = p_dist_plot,
      balance_plot = balance_plot,
      summary = balance_summary
    )
  }

@@

# Define a list of matching methods and parameters, each with a unique name
methods <- list(
  list(name = "nearest_glm", method = "nearest", distance = "glm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_gam", method = "nearest", distance = "gam", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_gbm", method = "nearest", distance = "gbm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_lasso", method = "nearest", distance = "lasso", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_rpart", method = "nearest", distance = "rpart", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "subclass_glm", method = "subclass", distance = "glm", subclasses = 5, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_mahalanobis", method = "nearest", distance = "mahalanobis", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "cbps", use="weightit", method="cbps", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "entropy", use="weightit", method="ebal", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "bart", use="weightit", method="bart", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "twang_gbm", use="twanng", method = "twang", estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE)
)

@@

```{r diag_plots}

for (method_name in method_names) {
  # Get balance and common support plots for the method
  method_results <- results[[method_name]]
  love_plot <- method_results$love_plot
  balance_plot <- method_results$balance_plot
  p_dist_plot <- method_results$p_dist_plot

@@

  # Output section header for balanace plot of pscore (if exists)
  if (!is.null(balance_plot)) {
    cat("### Balance Plot of distance measure for", method_name, "\n\n")
    print(balance_plot)
    cat("\n\n")
  }

    # Output section header for box plot of p-scores (if exists)
  if (!is.null(p_dist_plot)) {
    cat("### Balance Plot of distance measure for", method_name, "\n\n")
    print(p_dist_plot)
    cat("\n\n")
  }
}
```

```{r save_diag_plots}

if(! dir.exists("ps_diagnostics")) {
  dir.create("ps_diagnostics")
}

for (method_name in method_names) {
  # Get balance and common support plots for the method
  print(paste0("saving plots for ", method_name))
  method_results <- results[[method_name]]
  method_result <- method_results$result

  if(method_name == "subclass_glm") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)

    # jitter plot
    png(file.path("ps_diagnostics", paste0(method_name, "_p_dist_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_results$p_dist_plot, type="jitter", interactive=FALSE)
    dev.off()
  } else if(method_name == "nearest_mahalanobis") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)
  } else if(class(method_results$result) == "matchit") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)

    # jitter plot
    png(file.path("ps_diagnostics", paste0(method_name, "_p_dist_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_results$p_dist_plot, type="jitter", interactive=FALSE)
    dev.off()
  } else if(class(method_results$result) == "ps") {
    # love plot
    png(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_result, plots = 3)
    dev.off()

    # balance plot
    ps_data <- method_result$data
    ps_data$p_score <- method_result$ps$es.mean.ATT

    bplot <- ggplot(ps_data, aes(x=p_score, fill=person_ever_lapse)) + 
      geom_density(alpha=0.4)
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), bplot, dpi=300)

    # jitter plot
    png(file.path("ps_diagnostics", paste0(method_name, "_p_dist_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_result, plots = 2)
    dev.off()
  } else if(class(method_result) == "weightit") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)
  }
}
```
